---
title: ECON494 - Homework 2
author: Berke Ayyildizli - Melih Kiltoprak - Demirhan Izer - Aysegul Guler
---

1.  Run the following code to install or load (if already installed) the pacman package, and then install or load (if already installed) the listed package inside the pacman::p_load() function.

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  sf,       # vector data operations
  dplyr,    # data wrangling
  ggplot2,  # for map creation
  modelsummary, # regression table generation
  patchwork # arrange multiple plots
)

```

Run the following code to define the theme for maps:

```{r}
theme_for_map <- 
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_line(color = "transparent"),
    panel.background = element_blank(),
    plot.background = element_rect(fill = "transparent", color = "transparent")
  )

```

•Read in trial_design.rds .

```{r}
library(sf)

trial_design <- readRDS("HMW2_data/trial_design.rds")
str(trial_design)
```

•The above dataset contains the nitrogen rates across aerial units. Please create a map showing the spatial distribution of the variable NRATE .

```{r}
library(ggplot2)


custom_palette <- c("#fef0d9", "#fdcc8a", "#fc8d59", "#e34a33", "#b30000")

ggplot(trial_design) +
  geom_sf(aes(fill = NRATE)) +
  scale_fill_gradientn(colors = custom_palette) +
  ggtitle("Spatial Distribution of NRATE") +
  theme_for_map


```

•Now, read in the information concerning the “as applied nitrogen” (NH3.gpkg), “electric conductivity” (ec.shp), and “collected yield.” (yield.rds) . To do so, you will have to eomply st_read() and readRDS .

```{r}
library(sf)

nh3_data <- st_read("HMW2_data/NH3.gpkg")
ec_data <- st_read("HMW2_data/ec.shp")
yield_data <- readRDS("HMW2_data/yield.rds")

str(nh3_data)
str(ec_data)
str(yield_data)

```

•Create a figure showing the spatial distribution of the three variables. First create a map of each variable, and then combine them into one figure using the patchwork package

–To stack the figures vertically, label each individual figure as g_yield, g_NH3, and g_ec.

–Then write: g_yield / g_NH3 / g_ec

```{r}
library(ggplot2)
library(sf)
library(patchwork)


nh3_data <- st_read("HMW2_data/NH3.gpkg")
ec_data <- st_read("HMW2_data/ec.shp")
yield_data <- readRDS("HMW2_data/yield.rds")

color_palette <- c("#ffffcc", "#ffeda0", "#fed976", "#feb24c", "#fd8d3c", "#fc4e2a", "#e31a1c", "#bd0026", "#800026")


g_yield <- ggplot(yield_data) +
  geom_sf(aes(color = yield)) +
  scale_color_gradientn(colors = color_palette) +
  ggtitle("Spatial Distribution of Yield") +
  theme_for_map

g_NH3 <- ggplot(nh3_data) +
  geom_sf(aes(color = aa_NH3)) +
  scale_color_gradientn(colors = color_palette) +
  ggtitle("Spatial Distribution of NH3") +
  theme_for_map

g_ec <- ggplot(ec_data) +
  geom_sf(aes(color = ec)) +
  scale_color_gradientn(colors = color_palette) +
  ggtitle("Spatial Distribution of EC") +
  theme_for_map

combined_plot <- g_yield / g_NH3 / g_ec

print(combined_plot)

```

•Instead of using plot as the observation unit (the term “plot” refers to a piece of land), we would like to create subplots inside each of the plots and make them the unit of analysis. This will avoid hiding the within-plot spatial heterogeneity of EC. Please divide each plot into six subplots.

•Create a figure displaying the subplots.

```{r}
library(sf)
library(dplyr)
library(ggplot2)

create_subplots <- function(plot_geometry) {
  plot_bbox <- st_bbox(plot_geometry)
  width <- (plot_bbox$xmax - plot_bbox$xmin) / 3
  height <- (plot_bbox$ymax - plot_bbox$ymin) / 2
  
  subplots <- list()
  for (i in 0:2) {
    for (j in 0:1) {
      xmin <- plot_bbox$xmin + i * width
      xmax <- xmin + width
      ymin <- plot_bbox$ymin + j * height
      ymax <- ymin + height
      subplot <- st_polygon(list(rbind(
        c(xmin, ymin),
        c(xmin, ymax),
        c(xmax, ymax),
        c(xmax, ymin),
        c(xmin, ymin)
      )))
      subplots <- c(subplots, st_sfc(subplot, crs = st_crs(plot_geometry)))
    }
  }
  return(subplots)
}

subplots_list <- do.call(c, lapply(st_geometry(trial_design), create_subplots))
subplots_sf <- st_sf(geometry = st_sfc(subplots_list, crs = st_crs(trial_design)))

ggplot(subplots_sf) +
  geom_sf() +
  ggtitle("Subplots within Each Plot") +
  theme_for_map

```

•Now identify the mean value of corn yield, nitrogen rate, and EC for each of the subplots using sf::aggregate() and sf::st_join().

```{r}
library(sf)
library(dplyr)
library(ggplot2)

create_subplots <- function(plot_geometry) {
  plot_bbox <- st_bbox(plot_geometry)
  width <- (plot_bbox$xmax - plot_bbox$xmin) / 3
  height <- (plot_bbox$ymax - plot_bbox$ymin) / 2
  
  subplots <- list()
  for (i in 0:2) {
    for (j in 0:1) {
      xmin <- plot_bbox$xmin + i * width
      xmax <- xmin + width
      ymin <- plot_bbox$ymin + j * height
      ymax <- ymin + height
      subplot <- st_polygon(list(rbind(
        c(xmin, ymin),
        c(xmin, ymax),
        c(xmax, ymax),
        c(xmax, ymin),
        c(xmin, ymin)
      )))
      subplots <- c(subplots, st_sfc(subplot, crs = st_crs(plot_geometry)))
    }
  }
  return(subplots)
}

subplots_list <- do.call(c, lapply(st_geometry(trial_design), create_subplots))
subplots_sf <- st_sf(geometry = st_sfc(subplots_list, crs = st_crs(trial_design)))

subplots_sf$id <- 1:nrow(subplots_sf)

yield_subplots <- st_join(subplots_sf, yield_data)

nh3_subplots <- st_join(subplots_sf, nh3_data)

ec_subplots <- st_join(subplots_sf, ec_data)

mean_yield <- aggregate(yield_subplots["yield"], by = list(yield_subplots$id), FUN = mean)
mean_yield <- st_as_sf(mean_yield, geometry = subplots_sf$geometry[mean_yield$Group.1])
mean_nh3 <- aggregate(nh3_subplots["aa_NH3"], by = list(nh3_subplots$id), FUN = mean)
mean_nh3 <- st_as_sf(mean_nh3, geometry = subplots_sf$geometry[mean_nh3$Group.1])
mean_ec <- aggregate(ec_subplots["ec"], by = list(ec_subplots$id), FUN = mean)
mean_ec <- st_as_sf(mean_ec, geometry = subplots_sf$geometry[mean_ec$Group.1])

mean_values <- subplots_sf %>%
  st_join(mean_yield, join = st_within) %>%
  st_join(mean_nh3, join = st_within) %>%
  st_join(mean_ec, join = st_within)

print(mean_values)


```

•Provide a visualization of the subplot-level data.

```{r}
library(ggplot2)
library(patchwork)

custom_palette <- c("#fef0d9", "#fdcc8a", "#fc8d59", "#e34a33", "#b30000")

g_mean_yield <- ggplot(mean_values) +
  geom_sf(aes(fill = yield)) +
  scale_fill_gradientn(colors = custom_palette) +
  ggtitle("Mean Yield per Subplot") +
  theme_for_map

g_mean_nh3 <- ggplot(mean_values) +
  geom_sf(aes(fill = aa_NH3)) +
  scale_fill_gradientn(colors = custom_palette) +
  ggtitle("Mean NH3 per Subplot") +
  theme_for_map

g_mean_ec <- ggplot(mean_values) +
  geom_sf(aes(fill = ec)) +
  scale_fill_gradientn(colors = custom_palette) +
  ggtitle("Mean EC per Subplot") +
  theme_for_map

combined_plot <- g_mean_yield / g_mean_nh3 / g_mean_ec

print(combined_plot)

```

•Now you should be able to run a regression model.

```{r}
library(dplyr)
library(ggplot2)

custom_palette <- c("#fef0d9", "#fdcc8a", "#fc8d59", "#e34a33", "#b30000")

regression_data <- mean_values %>%
  filter(!is.na(yield) & !is.na(aa_NH3) & !is.na(ec))

model <- lm(yield ~ aa_NH3 + ec, data = regression_data)

summary(model)

plot_aa_NH3 <- ggplot(regression_data, aes(x = aa_NH3, y = yield)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  ggtitle("Yield vs. NH3") +
  theme_minimal()

plot_ec <- ggplot(regression_data, aes(x = ec, y = yield)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  ggtitle("Yield vs. EC") +
  theme_minimal()

print(plot_aa_NH3)
print(plot_ec)

```
